---
title: "Data Wrangling Coding Challenge"
author: "Dustyn Lewis"
date: "2025-03-25"
output:
  pdf_document:
    keep_md: yes
---
Load the required libraries

```{r}
library(tidyverse)
getwd()
```

1) Read in the CSV files

```{r}
diversity <- read.csv("DiversityData.csv")
metadata <- read.csv("MetaData.csv")
```

2) Join the two data frames by common column "code"

```{r}
alpha <- left_join(diversity, metadata, by = "Code")
```

3) Calculate Pielou's eveness index (*Shannon*/log(*Richness*))

```{r}
alpha_even <- alpha %>%
  mutate(even = shannon/log(richness))
```

4) Summarize mean and standard error of evenness grouped by *Crop* over *Time_Point*

```{r}
alpha_average <- alpha_even %>%
  group_by(Crop, Time_Point) %>%
  summarize(  
    mean.even = mean(even, na.rm = TRUE), 
    n=n(), 
    sd.even = sd(even, na.rm = TRUE), 
    se.even = sd.even/sqrt(n) ) %>%
  print()
```

5) Calculate differences between crops (Soil - Cotton, Soil - Soybean) for evenness

```{r}
alpha_average2 <- alpha_average %>%
  select(Crop:mean.even) %>%
  pivot_wider(names_from = Crop, values_from = mean.even) %>%
  mutate(diff.cotton.even = Soil - Cotton, diff.soybean.even = Soil - Soybean) %>%
  print()
```

6) Creating plots to compare differences over *Time_Point*

```{r}
alpha_long <- alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%
  pivot_longer( cols = c(diff.cotton.even, diff.soybean.even), 
               names_to = "diff", values_to = "values" ) 
  ggplot(alpha_long, aes(x = Time_Point, y = values, color = diff)) +
    geom_line() +
    labs( 
      title = "Difference in Evenness Over Time", 
      x = "Time Point", 
      y = "Difference in Evenness"
      )
```

7) Here is a [clickable link to GitHub](https://github.com/Dustyn-T-Lewis/Coding-Challenge-5)